Commutative character sheaves and geometric types for supercuspidal representations

### 2017-05-07 Geometric types.tex
Some work on descent

In order to study commutative character sheaves, we need some auxiliary categories.
\begin{itemize}
\item Let $\Loc(G)$ and $\Loc(G\ab)$ be the categories of local systems on $G$ and $G\ab$, respectively.
\item Let $\Loc_{G\der}(G)$ be the category of $G\der$-equivariant local systems on $G$, whose definition we now recall.
Let $n : G\der\times G\to G$ be the restriction of $m : G\times G\to G$ to $G\der\times G$,
let $p : G\der\times G \to G$ be projection to the second component, and let $s: G \to G\der\times G$ be given by $s(g) = (1,g)$.
Objects in $\Loc_{G\der}(G)$ are pairs $(\cs{L},\nu)$ where $\cs{L}$ is a local system on $G$ and
$\nu : n^*\cs{L} \to p^*\cs{L}$ is an isomorphism such that the following two diagrams commute.
\[
\begin{tikzcd}
s^*n^*\cs{L} \arrow{r}{s^*(\nu)} & s^*p^*\cs{L}\\
\cs{L} \arrow[equal]{u}{n\circ s = \id_G}  \arrow{r}{\id_{\cs{L}}} & \cs{L} \arrow[equal]{u}[swap]{p\circ s = \id_G},
\end{tikzcd}
\]
and with $n_1(h_1,h_2,g) = (h_1h_2,g)$, $n_2(h_1,h_2,g) = (h_1,n(h_2,g))$ and $p_2(h_1,h_2,g) = (h_2,g)$,
\[
\begin{tikzcd}
& \arrow[equal]{dl}{n\circ n_2=n\circ n_1} n_2^*n^*\cs{L} \arrow{rr}{n_2^*(\nu)} && n_2^*p^*\cs{L} \arrow[equal]{dr}[swap]{p \circ n_2 = n\circ p_2}& \\
n_1^*n^*\cs{L} \arrow{dr}[swap]{n_1^*(\nu)} &&&& \arrow{dl}{p_2^*(\nu)} p_2^*n^*\cs{L}\\
& n_1^*p^*\cs{L} \arrow[equal]{rr}{p\circ n_1 = p\circ p_2} && p_2^*p^*\cs{L} .
\end{tikzcd}
\]
Morphisms are morphisms in $\Loc(G)$ commuting with $\nu$.
\item Let $\CSab'(G)$ be the category of triples $(\cs{L},\nu,\mu)$ where $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\nu)\in \Loc_{G\der}(G)$; thus, $\mu : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ and $\nu :  n^*\cs{L} \to p^*\cs{L}$ are isomorphisms.
Morphisms in $\Loc_{G\der}(G)$ are morphisms in $\Loc(G)$ commuting with $\mu$ and $\nu$. 
\end{itemize}

@@

We may now relate $\CSab(G)$ to $\CS(G\ab)$ via the quotient map $q : G \to G\ab$.

%%%%%%%%
\iffalse

\todo{Define $q\ab^* : \CS(G\ab) \to \CSab(G)$; missing definition of $\beta$ in the following commented section}
Now suppose $q : G \to Q$ is a regular epimorphism of smooth group schemes with kernel pair $(a,p)$
\[
\begin{tikzcd}
H\times G
 \arrow[shift left=2]{r}{a}
  \arrow[shift right=2,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
Q.
\end{tikzcd}
\]
Let $(\cs{L}_0, \mu_0)$ be a linear character sheaf on $Q$ and set $(\cs{L}, \mu) \ceq (q^*\cs{L}_0, (q\times q)^* \mu_0)$.
Consider the functor
\[
q^* : \CS(Q) \to \CS(G)
\]
given on objects by $(\cs{L}_0,\mu_0) \mapsto (\cs{L}, \mu)$; see \cite{cunningham-roe:13a}*{Lem 1.4}.
To see that $(q^*\cs{L}_0, (q\times q)^* \mu_0)$ is indeed a linear character sheaf on $G$,
verify \cite{cunningham-roe:13a}, CS.3], arguing as in \cite{cunningham-roe:13a}].
The linear character sheaf $(\cs{L},\mu)$ on $G$ comes equipped with a canonical isomorphism $\nu : a^* \cs{L} \to p^* \cs{L}$ defined by the following diagram of isomorphisms.
\[
\begin{tikzcd}[column sep=40]
a^*\cs{L} \arrow[equal]{d} \arrow[dashed]{r}{\nu} &  \arrow[equal]{d} p^*\cs{L} \\
a^* ( q^*\cs{L}_0) \arrow{r}{q\circ a = q\circ p}
& p^* (q^*\cs{L}_0)
\end{tikzcd}
\]
Since this isomorphism satisfies \eqref{ECS1}, it follows that $(\cs{L}, \mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$.
If $\alpha_0 : (\cs{L}_0,\mu_0) \to (\cs{L}_0',\mu_0')$ is a morphism in $\CS(Q)$, 
then $q^*(\alpha_0) : (\cs{L},\mu) \to (\cs{L}',\mu')$ satisfies \cite{cunningham-roe:13a}*{CS4}, so $\alpha$ is a morphism in $\CS(G)$.
These simple observations define the comparison functor
\[
q_H^* : \CS(Q) \to CS_H(G)
\]
and show that the functor $q^* : \CS(Q) \to \CS(G)$ factors according to the following commuting diagram of functors
\begin{equation}\label{qH}
\begin{tikzcd}
\CS(G) &\arrow{l}[swap]{q^*} 
\CS(Q) \arrow{dl}{q_H^*}\\
\arrow{u}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\end{equation}
The definition of $q_H^* : \CS(Q)\to \CS_H(G)$ will be revisited in the proof of the following result.

\begin{proposition} \label{prop:Gder_triv}
Suppose $G$ is a smooth group scheme and $(\cs{L},\mu) \in \CS(G)$.
Then the restriction of $\cs{L}$ to $G\der$ is trivial if and only if $(\cs{L},\mu) \cong q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G)$, for some $(\cs{L}\ab,\mu\ab) \in \CS(G\ab)$.
\end{proposition}
Proposition~\ref{prop:Gder_triv} will be proven in Section \ref{ssec:proof1}.

\fi
%%%%%%%%%%%%%%%%%

@@

Let $i\der : G\der \to G$ be the kernel of $q:G \to G\ab$ and define $j : G\der \to G\der\times G$ by $j(h) = (i\der(h),1_G)$. 
For $(\cs{L},\mu,\nu)\in \CSab'(G)$, $j^*(\nu) : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism. \todo{For $j^*(\nu)$ to make sense, need $j : G\der \to G\der \times G$}
This defines a functor
\[
\begin{array}{rcl}
\CSab'(G) &\to& \CSab(G) \\
(\cs{L},\mu,\nu) &\mapsto& (\cs{L},\mu,j^*(\nu)).
\end{array}
\]
Define $k: G\der\times G\to G\times G$ by $k(h,g) = (i\der(h),g)$.
Then for $(\cs{L},\mu,\beta) \in \CSab(G)$, define $\nu : n^*\cs{L} \to p^*\cs{L}$ 
%\todo{I assume this should be $p^*\cs{L}$?} by
\[
\begin{tikzcd}
& n^* \cs{L} \arrow{dl}{m\circ k = n} \arrow{r} & p^*\cs{L} &  \\ 
k^* m^* \cs{L} \arrow{dr}{k^*(\mu)} &&& (\EE)_{G\der} \boxtimes \cs{L} \arrow[equal]{ul} \\
& k^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & \cs{L}\vert_{G\der}\boxtimes \cs{L} \arrow{ur}{\beta\times 1}
\end{tikzcd}
\]
Then $(\cs{L},\mu,\nu)\in \CSab'(G)$ maps to $(\cs{L},\mu,\beta)$.
It now follows that $\CSab'(G) \to \CSab(G)$ is an equivalence. \todo{Is it easy to see why the morphisms that commute with $\mu$ and $\beta$ are the same as those commuting with $\mu$ and $\nu$?  Is this straight from the definition of the functor?}
\end{proof}

>>>>

In order to study commutative character sheaves, we need some auxiliary categories.
%\begin{itemize}
%\item 
Let $\Loc(G)$ and $\Loc(G\ab)$ be the categories of local systems on $G$ and $G\ab$, respectively.
%\item 
Let $\Loc_{G\der}(G)$ be the category of $G\der$-equivariant local systems on $G$, whose definition we now recall.
Let $n : G\der\times G\to G$ be the restriction of $m : G\times G\to G$ to $G\der\times G$,
let $p : G\der\times G \to G$ be projection to the second component, and let $s: G \to G\der\times G$ be given by $s(g) = (1,g)$.
Then $q : G \to G\ab$ is a regular epimorphism of smooth group schemes with kernel pair $(n,p)$.
\[
\begin{tikzcd}
G\der\times G
 \arrow[shift left=1]{r}{n}
  \arrow[shift right=1,swap]{r}{p}
&
G 
\arrow{r}{q}
& 
G\ab
\end{tikzcd}
\]
Objects in the category $\Loc_{G\der}(G)$ of $G\der$-equivariant local systems on $G$ are pairs $(\cs{L},\nu)$ where $\cs{L}$ is a local system on $G$ and $\nu : n^*\cs{L} \to p^*\cs{L}$ is an isomorphism such that the diagram
\[
\begin{tikzcd}
s^*n^*\cs{L} \arrow{r}{s^*(\nu)} & s^*p^*\cs{L}\\
\cs{L} \arrow[equal]{u}{n\circ s = \id_G}  \arrow{r}{\id_{\cs{L}}} & \cs{L} \arrow[equal]{u}[swap]{p\circ s = \id_G},
\end{tikzcd}
\]
commutes, and also such that the following diagram commutes, where $n_1(h_1,h_2,g) = (h_1h_2,g)$, $n_2(h_1,h_2,g) = (h_1,n(h_2,g))$ and $p_2(h_1,h_2,g) = (h_2,g)$.
\[
\begin{tikzcd}
& \arrow[equal]{dl}{n\circ n_2=n\circ n_1} n_2^*n^*\cs{L} \arrow{rr}{n_2^*(\nu)} && n_2^*p^*\cs{L} \arrow[equal]{dr}[swap]{p \circ n_2 = n\circ p_2}& \\
n_1^*n^*\cs{L} \arrow{dr}[swap]{n_1^*(\nu)} &&&& \arrow{dl}{p_2^*(\nu)} p_2^*n^*\cs{L}\\
& n_1^*p^*\cs{L} \arrow[equal]{rr}{p\circ n_1 = p\circ p_2} && p_2^*p^*\cs{L} .
\end{tikzcd}
\]
Morphisms in $\Loc_{G\der}(G)$ are morphisms in $\Loc(G)$ commuting with $\nu$.

%\item 
Let $\CS_{G\der}(G)$ be the category of triples $(\cs{L},\nu,\mu)$ where $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\nu)\in \Loc_{G\der}(G)$.
%Thus, $\mu : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ and $\nu :  n^*\cs{L} \to p^*\cs{L}$ are isomorphisms.
Morphisms in $\CS_{G\der}(G)$ are morphisms in $\CS(G)$ commuting with $\nu$. 
%\end{itemize}

%%%%%%%%
%\iffalse

@@

\todo{Define $q\ab^* : \CS(G\ab) \to \CSab(G)$; missing definition of $\beta$ in the following commented section}

%Let $(\cs{L}_0, \mu_0)$ be a linear character sheaf on $Q$ and set $(\cs{L}, \mu) \ceq (q^*\cs{L}_0, (q\times q)^* \mu_0)$.
Consider the functor
\[
q^* : \CS(G\ab) \to \CS(G)
\]
given on objects by $(\cs{L}\ab,\mu\ab) \mapsto (q^*\cs{L}\ab, (q^2)^* \mu\ab)$ where $q^2= q\times q$; see \cite{cunningham-roe:13a}*{Lem 1.4}.
To see that $(q^*\cs{L}\ab, (q^2)^* \mu\ab)$ is indeed a linear character sheaf on $G$,
verify \cite{cunningham-roe:13a}*{CS.3}. %, arguing as in \cite{cunningham-roe:13a}.
The linear character sheaf $(\cs{L},\mu)$ on $G$ comes equipped with a canonical isomorphism $\nu : n^* \cs{L} \to p^* \cs{L}$ defined by the following diagram of isomorphisms.
\[
\begin{tikzcd}[column sep=40]
n^*\cs{L} \arrow[equal]{d} \arrow[dashed]{r}{\nu} &  \arrow[equal]{d} p^*\cs{L} \\
n^* ( q^*\cs{L}\ab) \arrow{r}{q\circ n = q\circ p}
& p^* (q^*\cs{L}\ab)
\end{tikzcd}
\]
Then $(\cs{L}, \mu,\nu)$ is an object in $\CS_{G\der}(G)$.
If $\alpha_0 : (\cs{L}_0,\mu_0) \to (\cs{L}_0',\mu_0')$ is a morphism in $\CS(Q)$, 
then $q^*(\alpha_0) : (\cs{L},\mu) \to (\cs{L}',\mu')$ satisfies \cite{cunningham-roe:13a}*{CS4}, so $\alpha$ is a morphism in $\CS(G)$.
These simple observations define the comparison functor
\[
q_H^* : \CS(Q) \to CS_H(G)
\]
and show that the functor $q^* : \CS(Q) \to \CS(G)$ factors according to the following commuting diagram of functors
\begin{equation}\label{qH}
\begin{tikzcd}
\CS(G) &\arrow{l}[swap]{q^*} 
\CS(Q) \arrow{dl}{q_H^*}\\
\arrow{u}{\text{forget}} \CS_{H}(G). & 
\end{tikzcd}
\end{equation}
The definition of $q_H^* : \CS(Q)\to \CS_H(G)$ will be revisited in the proof of the following result.

We may now begin to relate $\CSab(G)$ to $\CS(G\ab)$ via the quotient map $q : G \to G\ab$.

%\fi
%%%%%%%%%%%%%%%%%

@@

Let $i\der : G\der \to G$ be the kernel of $q:G \to G\ab$ and define $j : G\der \to G\der\times G$ by $j(h) = (h,1_G)$. 
For $(\cs{L},\mu,\nu)\in \CSab'(G)$, $j^*(\nu) : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ is an isomorphism. \todo{For $j^*(\nu)$ to make sense, need $j : G\der \to G\der \times G$}
This defines a functor
\[
\begin{array}{rcl}
\CSab'(G) &\to& \CSab(G) \\
(\cs{L},\mu,\nu) &\mapsto& (\cs{L},\mu,j^*(\nu)).
\end{array}
\]
Define $k: G\der\times G\to G\times G$ by $k(h,g) = (i\der(h),g)$.
Then for $(\cs{L},\mu,\beta) \in \CSab(G)$, define $\nu : n^*\cs{L} \to p^*\cs{L}$ 
%\todo{I assume this should be $p^*\cs{L}$?} by
\[
\begin{tikzcd}
& n^* \cs{L} \arrow{dl}{m\circ k = n} \arrow{r} & p^*\cs{L} &  \\ 
k^* m^* \cs{L} \arrow{dr}{k^*(\mu)} &&& (\EE)_{G\der} \boxtimes \cs{L} \arrow[equal]{ul} \\
& k^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & \cs{L}\vert_{G\der}\boxtimes \cs{L} \arrow{ur}{\beta\times 1}
\end{tikzcd}
\]
Then $(\cs{L},\mu,\nu)\in \CSab'(G)$ maps to $(\cs{L},\mu,\beta)$.
It now follows that $\CSab'(G) \to \CSab(G)$ is an equivalence. \todo{Is it easy to see why the morphisms that commute with $\mu$ and $\beta$ are the same as those commuting with $\mu$ and $\nu$?  Is this straight from the definition of the functor?}
\end{proof}
